Analysis-ready VCF at Biobank scale using Zarr

### 2025-01-23 paper.tex
Update second GPU section

To demonstrate the ease and efficiency with this data stored in Zarr 
format can be processed on a GPU, we developed two proof-of-concept
applications. In the first we re-implement the af-dist calculation
used throughout as an example of a non-trivial whole-matrix operation
on GPU. In the second, we demonstrate some emerging technology 
in which Zarr data can be decompressed directly in GPU memory. 
In both cases we demonstrate on chromosome 22 of 
1000 Genomes WGS data~\citep{byrska2022high}. This consists
of 26,504 variants and 2504 samples [and more details about the zarr].
Benchmarks in this section were performed on a
machine with an Intel Xeon Silver 4216 processor,
Nvidia A6000 RTX GPU and NVMe SSD.

@@

Chunk decompression comprises a significant fraction of processing 
time [FIXME crossrefs?]

>>>>

To demonstrate the ease and efficiency with this data stored in Zarr 
format can be processed on a GPU, we developed two proof-of-concept
applications. In the first we re-implement the af-dist calculation
used throughout as an example of a non-trivial whole-matrix operation
on GPU. In the second, we demonstrate some emerging technology 
in which Zarr data can be decompressed directly in GPU memory. 
In both cases we demonstrate on chromosome 22 of 
1000 Genomes WGS data~\citep{byrska2022high}. This consists
of 96,514 variants and 2504 samples [and more details about the zarr].
Benchmarks in this section were performed on a
machine with an Intel Xeon Silver 4216 processor,
Nvidia A6000 RTX GPU and NVMe SSD.

@@

Chunk decompression comprises a significant fraction of processing 
time [FIXME crossrefs?] and host-to-GPU data transfer is 
often a bottleneck. It is therefore useful in some applications 
to directly decompress Zarr chunks into GPU memory. 
KvikIO is a Python package that provides access to optimised
GPU IO routines, and providing support for Zarr along with
access to several GPU optimised compression codecs.
To explore the performance of these codecs we recompressed
the \texttt{call\_AD} (allele depth) array
(96514, 2504, 7) [int16 more details]. In call cases it took around 1s 
to compress and store this 3.15GiB array. [Details about compression codecs?]
We then measured the time to decompress these arrays and 
copy back to host memory [Why??]. This required about 0.6s with the 
fastest codecs. For reference, using the Zstandard codec on a single
CPU thread required about 7.6s.
